%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                        %
%    TypeRex OCaml Studio                                                %
%      Thomas Gazagnaire, Fabrice Le Fessant                             %
%                                                                        %
%    OCaml                                                               %
%      Xavier Leroy, projet Cristal, INRIA Rocquencourt                  %
%                                                                        %
%  Copyright 2011-2012 OCamlPro                                          %
%  Copyright 1996-2011 INRIA.                                            %
%  All rights reserved.  This file is distributed under the terms of     %
%  the GNU Public License version 3.0.                                   %
%                                                                        %
%  TypeRex is distributed in the hope that it will be useful,            %
%  but WITHOUT ANY WARRANTY; without even the implied warranty of        %
%  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the         %
%  GNU General Public License for more details.                          %
%                                                                        %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{ocp-type}

The \verb!ocp-type! command-line tool is the ``type-only'' OCaml compiler
which is used to extract the binary annotations needed by the
\typerex\ environment. It may be invoked directly or through the
\verb!ocp-wrapper! tools (see below). It accepts the same options and
arguments as \verb!ocamlc!, and the specific option
\verb!-save-types!, to actually write the binary data to a file.

\subsubsection*{Pre-processing}
\verb!ocp-type! also accepts pre-processors through the \verb!-pp!
option, which should output either
\begin{itemize}
\item an OCaml source file (if possible with line-number directives to
  allow an accurate use of \typerex), or
\item an OCaml dumped AST of one of the supported versions (this is the
  default behavior of \verb!camlp4!), or
\item a Camlp4 dumped AST of one of the supported versions, which can
  be achieved by passing the option \verb!-printer Camlp4AstDumper! to
  \verb!camlp4! (this is done automatically if you use
  \verb!ocp-wrapper!).
\end{itemize}
Note that the third option is better than the second one with respect
to locations (the accuracy of which is instrumental to \typerex\
working).

\subsubsection*{Libraries}

\verb!ocp-type! accepts the special form
{\verbsize\begin{verbatim}
    ocp-type -save-types -a <units> -o <target>
\end{verbatim}}
where
\begin{itemize}
\item the argument units may be \verb!cmi!, \verb!cmo!,  \verb!cmti!, \verb!cmt!, \verb!mli!, or  \verb!ml! files, and
\item the target is ignored.
\end{itemize}
Only arguments ending in \verb!ml! or \verb!mli! will be considered,
and typed into \verb!cmt! or \verb!cmti! files.

\subsubsection*{Module packs}

\verb!ocp-type! accepts the special form
{\verbsize\begin{verbatim}
    ocp-type -save-types -pack <units> -o <target>
\end{verbatim}}
where
\begin{itemize}
\item the argument units may be \verb!cmi!, \verb!cmo!,  \verb!cmti!, \verb!cmt!, \verb!mli!, or  \verb!ml! files, and
\item the target may be a \verb!cmo! or \verb!cmt! file.
\end{itemize}
Howerver, input will always look at \verb!cmt! or \verb!cmti! (or
\verb!cmi! as a fallback) files (possibly generating them if \verb!ml!
or \verb!mli!  are given), and the output will always be written in a
\verb!cmt! file.  If a \verb!mli! file exists for the name of the
pack, then \verb!ocp-type! will look for a compiled interface file for
it in \verb!cmti! format (or \verb!cmi! as a fallback) and match the
result of packing the arguments against this signature.
