%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                        %
%                        TypeRex OCaml Studio                            %
%                                                                        %
%                           Tiphaine Turpin                              %
%                                                                        %
%  Copyright 2011-2012 INRIA Saclay - Ile-de-France / OCamlPro           %
%  All rights reserved.  This file is distributed under the terms of     %
%  the GNU Public License version 3.0.                                   %
%                                                                        %
%  TypeRex is distributed in the hope that it will be useful,            %
%  but WITHOUT ANY WARRANTY; without even the implied warranty of        %
%  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the         %
%  GNU General Public License for more details.                          %
%                                                                        %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{ocp-wrapper}

The command \verb!ocp-wrapper! and the specialized commands
\verb!ocp-ocamlc!, \verb!ocp-ocamlopt!, \verb!ocp-ocamlc.opt!, and
\verb!ocp-ocamlopt.opt! simplify the generation of binary annotations
by invoking \verb!ocp-type! with the appropriate options, as part of
the usual compilation commands. Calling
{\verbsize \begin{verbatim}
        ocp-wrapper -save-types <command> <options and arguments>
\end{verbatim}}
\noindent where \verb!<command>! is one of the OCaml compilers first
invokes this compiler with the exact same options and arguments, and
then runs \verb!ocp-type! (unless the command-line was a linking-only
phase) with the right options and arguments which are deduced from the
original command ones.

\subsubsection*{Shortcuts}

The four \verb!ocp-*! commands are shortcuts for
\verb!ocp-wrapper -save-types <command>! which are useful when a
single executable program is required as compiling command.

\subsubsection*{Options}

\begin{itemize}
\item \verb!-with-ocp-type!, \verb!-with-ocamlc!, \ldots: allow to
  customize the \verb!ocp-type!, \verb!ocamlc!, \ldots commands which
  are run.

\item \verb!-v! print the \verb!ocp-type! command which is executed on
  stderr.
\end{itemize}

\subsubsection*{Pre-processing}
Any \verb!-pp! option appearing in the command line is transformed as
follows when passed to \verb!ocp-type!:
\begin{itemize}
\item if the pre-processor command is (a variant of) \verb!camlp4!,
  and unless option \verb!-no-wrap-camlp4! is passed to \verb!ocp-wrapper!,
  then the option \verb!-printer Camlp4AstDumper! is added, which
  yields more accurate location information,
\item otherwise, the command is left unchanged.
\end{itemize}

\subsubsection*{Libraries and Module Packs}

\verb!ocp-wrapper! accepts the special forms
{\verbsize\begin{verbatim}
    ocp-wrapper -save-types <compiler> -a <units> -o <target>
    ocp-wrapper -save-types <compiler> -pack <units> -o <target>
\end{verbatim}}
Arguments and targets ending in \verb!cmx! are converted into \verb!cmo!
and targets into \verb!cmt! as appropriate (see the documentation for
\verb!ocp-type!).
